---
title: "class09"
author: "Andres Rivero"
date: "10/27/2021"
output: html_document
---

```{r}
fna.data <- "WisconsinCancer.csv"

```


```{r}
wisc.df <- read.csv(fna.data, row.names=1)
head(wisc.df)
```

```{r}
wisc.data <- wisc.df[,-1]
diagnosis <- factor(wisc.df$diagnosis, levels = c("B", "M"))
print(diagnosis)
```
**Q1: How many observations are in this dataset?**
```{r}
dim(wisc.data)

```
It looks like there is 569 different observations.


**Q2: How many of the observations have a malignant diagnosis?**

```{r}
table(diagnosis)
```
There is 212 malignant observations

**Q3: How many variables/features in the data are suffixed with _mean?**
```{r}
length(grep("_mean", colnames(wisc.data)))
```
10 columns have the term "_mean" in them



Next: PCA

Check column means and standard deviations
```{r}
colMeans(wisc.data)
apply(wisc.data,2,sd)

wisc.pr <- prcomp(wisc.data, scale. = TRUE)

```
```{r}
summary(wisc.pr)
```

**Q4: From your results, what proportion of the original variance is captured by the first principal components (PC1)?**
44.27%

**Q5: How many principal components (PCs) are required to describe at least 70% of the original variance in the data?**
3

**Q6: How many principal components (PCs) are required to describe at least 90% of the original variance in the data?**
7

```{r}
biplot(wisc.pr)
```
**Q7: What stands out to you about this plot? Is it easy or difficult to understand? Why?**

Can't understand anything on the plot, impossible to read

```{r}
plot( wisc.pr$x , col = diagnosis , 
     xlab = "PC1", ylab = "PC2")
```
**Q8: Generate a similar plot for principal components 1 and 3. What do you notice about these plots?**

```{r}
  plot( wisc.pr$x[, c(1,3)] , col = diagnosis , 
     xlab = "PC1", ylab = "PC3")
```
There seems to be two different cluster of cells, red dots(or malignant) and black dots (or benign)


**Creating ggplot for our data**

```{r}
df <- as.data.frame(wisc.pr$x)
df$diagnosis <- diagnosis
```

```{r}
library(ggplot2)
ggplot(df) + aes(PC1, PC2, col=diagnosis) + 
  geom_point()
```
**Calculating the variance now**

```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)
```
```{r}
pve <- pr.var/(sum(pr.var))
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

```{r}
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```
**Q9: For the first principal component, what is the component of the loading vector (i.e. wisc.pr$rotation[,1]) for the feature concave.points_mean?**

```{r}
wisc.pr$rotation["concave.points_mean",1]

```

**Q10: What is the minimum number of principal components required to explain 80% of the variance of the data?**

```{r}
summary(wisc.pr)
```
We need The first 5 PCs to get to at least 80%




