---
title: "HW Class 06"
author: "Andres Rivero"
date: "10/16/2021"
output: pdf_document
---

**Can you improve this analysis code?**


library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug  
s2 <- read.pdb("1AKE") # kinase no drug  
s3 <- read.pdb("1E4Y") # kinase with drug  


s1.chainA <- trim.pdb(s1, chain="A", elety="CA")  
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")  
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")  


s1.b <- s1.chainA$atom$b  
s2.b <- s2.chainA$atom$b  
s3.b <- s3.chainA$atom$b  


plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")  
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")  
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")  


**We will start by getting a snippet of each of the different codes ran for the different enzymes. What does this code do?**

```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")

```


It seem like the code looks the name of a given protein on a database, downloads information about it into the R environment, trims some of that info and gives a plot for said protein. 

**How could we simplify/or generalize the code shown above to make our function?**  
We could start by substituting "x" in place of 

sx <- read.pdb(x) #This would get the info about "x" protein from the database  
sx.chainA <- trim.pdb(sx, chain="A", elety="CA") #This would trim that information  
sx.b <- sx.chainA$atom$b #This would select the part of all that information that we want to plot  
plotb3(sx.b, sse=s1.chainA, typ="l", ylab="Bfactor") #And this would plot it


We could include that in the body of our function and see if it manages to plot a protein (we will try with the third one)

```{r}
plotprot <- function(x){
  sx <- read.pdb(x) 
sx.chainA <- trim.pdb(sx, chain="A", elety="CA")
sx.b <- sx.chainA$atom$b
plotb3(sx.b, sse=s1.chainA, typ="l", ylab="Bfactor")
}
plotprot("1E4Y")
```

The function seemed to work.  
By just giving it the protein code it automatically gives us the same plot that we were getting with 4 different lines of code before.

We will do another test with the second protein from the example:

```{r}
plotprot("1AKE")

```


And a final test with the very first one, and compare it to the first graph obtained in this pdf:

```{r}
plotprot("4AKE")
```

It worked and the plots look the same. **SUCCESS**




