---
title: "Class06"
author: "Andres Rivero"
date: "10/15/2021"
output: pdf_document
---

## Quick Rmarkdown intro

We can write test of course just like any file. **We can style text to be bold** or *italic*.

Do:  
- this
- and that
- and another thing

This is more text  
and this is a new line


---

We can include some code:

```{r}
plot(1:10)
```
```{r}
#This is a comment and will not be passed to R
plot(cars)
```

```{r}
# Ctrl+ALt+I
```

## Time to write a function

> **Q1**. Write a function grade() to determine an overall grade from a vector of student homework
assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an
NA value) this can be used as a score to be potentially dropped. Your final function should be
adquately explained with code comments and be able to work on an example class gradebook
such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]


```{r}
# Example input vectors to start with

student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)

```
First I want to find the lowest score. I can yse the **min()** to find it and the **which.min() function to find where it is (i.e. its position in the vector).
```{r}

min(student1)
which.min(student1)
```

I can use minus to get everything in the vector but the lowest score.

```{r}
student1[-(which.min(student1))]
```

Now I can call the **mean()** function to get the average

```{r}
mean(student1[-(which.min(student1))])
```
```{r}
mean(student2[-(which.min(student2))])
```

NO!
But why not? 
```{r}
which.min(student2)
```
```{r}
mean(student2)
```
```{r}
mean(student2, na.rm=TRUE)
```
One great idea is to replace the NA values with zero.

```{r}
which(is.na(student2))

```
The is.na function returns a logical vector where TRUE elements indicate the presence
```{r}
is.na(student2)
```

Lets replace NA with 0
```{r}
student.prime <- student2
student.prime[is.na(student.prime)]=0
student.prime
```

Putting bits together

```{r}
student.prime <- student2
student.prime[is.na(student.prime)]=0
mean(student.prime[-which.min(student.prime)])
```
How about student 3
```{r}
student.prime <- student3
student.prime[is.na(student.prime)]=0
mean(student.prime[-which.min(student.prime)])
```
We got it to work, now we simplify it.

```{r}
x <- student3
x[is.na(x)]=0
mean(x[-which.min(x)])
```
```{r}
student4 <- c(100, NA, 90, "90", 90, 90, 97, 80)
```

```{r}
as.numeric(student4)
```
```{r}
x <- student4
x <- as.numeric(x)
x[is.na(x)]=0
mean(x[-which.min(x)])
```
Finally we can write our function:

All functions have at least 3 things. 

A name, input arg and a body

**Q1**
```{r}
grade <- function(x) {
  x <- as.numeric(x)
x[is.na(x)]=0
mean(x[-which.min(x)])
}
```

```{r}
grade(student1)
```
## For the whole class


To grade the whole class, we read the gradebook for the class.

```{r}
gradebook <- "https://tinyurl.com/gradeinput"
scores <- read.csv(gradebook, row.names = 1)
scores
```

Super useful function=  **apply()**. We can use it to grade all our students with our **grade()** function.

```{r}
apply(scores, 1, grade)
```

**Q2** Finding out the student with the highest score:

```{r}
which.max(apply(scores, 1, grade))
max(apply(scores, 1, grade))
```

**Q3** Finding out the toughest homework across the class:

```{r}
apply(scores, 2, mean )
```

```{r}
apply(scores, 2, mean, na.rm=TRUE)
```
Replace NA values to zero
```{r}

mask <- scores
mask[is.na(mask)]= 0
apply(mask,2, mean)
which.min(apply(mask,2, mean))
min(apply(mask,2, mean))
```







