---
title: "Halloween Candy"
author: "Andres Rivero"
date: "10/29/2021"
output: pdf_document
---
```{r}
candy_file <- "candy-data.txt"
candy = read.csv(candy_file, row.names=1)
head(candy)
```

```{r}
dim(candy)
```

**Q1**
85 different types of candy


```{r}
sum(candy$fruity)
```
**Q2**
38 of them



```{r}
candy["Kit Kat", ]$winpercent
```
**Q3**
Kit Kat, 76%

**Q4**
76%

**Q5**
```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```
49%

```{r}
#install.packages("skimr")
#library("skimr")
```
```{r}
#skim(candy)
```
**Q6**
Yes, the histogram column for example


**Q7**
0 means it doesn't have it or is not considered to be chocolate candy, and 1 means it is considered to have chocolate.


**Q8**
```{r}
hist(candy$winpercent)
```
**Q9**
No it isn't.

**Q10**
Below

**Q11**
```{r}
chocolate <- candy[as.logical(candy$chocolate),]$winpercent
mean(chocolate)
```
```{r}
fruity <- candy[as.logical(candy$fruity),]$winpercent
mean(fruity)
```
Chocolate is higher

**Q12**
```{r}
t.test(chocolate, fruity)
```

**Q13**
```{r}
head(candy[order(candy$winpercent),], n=5)
```

**Q14**
```{r}
tail(candy[order(candy$winpercent),], n=5)
```

**Q15**
```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```
**Q6**

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, reorder(row.names(candy), winpercent)) +
  geom_col()
```


```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```
**Q17**
Charleston Chew

**Q18**
Starburst

**Q19**

#install.packages("ggrepel")
#library("ggrepel")

#ggplot(candy) +
  #aes(winpercent, pricepercent, label=rownames(candy)) +
  #geom_point(col=my_cols) + 
  #geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)

Reeses Miniatures


**Q20**
```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```


```{r}
#install.packages("corrplot")
#library(corrplot)
#cij <- cor(candy)
#corrplot(cij)
```
**Q22**
Chocolate and Fruity

**Q23**
Chocolate and Bars


# Principle Component Analysis
```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```
